---
title: "hw01-shuai-shao.Rmd"
output: github_document
---
#2)
```{r}
library(readr)
col_types <- c(symboling = col_double(),
           normalized_losses = col_double(),
           make = col_character(), 
           fuel_type = col_character(), 
           aspiration = col_character(), 
           num_of_doors = col_character(),
           body_styles = col_character(),
           drive_wheels = col_character(),
           engine_location = col_character(), 
           wheel_base = col_double(), 
           length = col_double(), 
           width = col_double(), 
           height = col_double(), 
           curb_weight = col_integer(), 
           engine_type = col_character(), 
           num_of_cylinders = col_character(), 
           engine_size = col_integer(), 
           fuel_system = col_character(), 
           bore = col_double(), 
           stroke = col_double(), 
           compression_ratio = col_double(), 
           horsepower = col_integer(), 
           peak_rpm = col_integer(), 
           city_mpg = col_integer(), 
           highway_mpg = col_integer(), 
           price = col_integer())

auto <- read_csv('imports-85.data', col_types=col_types)
str(auto)
```

```{r}
names <- c('symboling', 'normalized_losses', 'make', 'fuel_type', 'aspiration', 
           'num_of_doors', 'body_styles', 'drive_wheels', 'engine_location', 
           'wheel_base', 'length', 'width', 'height', 'curb_weight', 
           'engine_type', 'num_of_cylinders', 'engine_size', 'fuel_system', 
           'bore', 'stroke', 'compression_ratio', 'horsepower', 'peak_rpm', 
           'city_mpg', 'highway_mpg', 'price')

colClasses <- c('double', 'double', rep('character',7),
             rep('double',4), 'integer', 'character', 'character', 'integer',
             'character', rep('double',3), rep('integer',5))
auto <- read.csv('imports-85.data', col.names=names, 
                 colClasses=colClasses, na.strings='?')
str(auto)
```

#3) 
# a. They take the values on the first row.
# For numerical values, they appear as X[value].
# For non-numerical values, they appear as [value]
# where [value] is the corresponding value on the first row.
# If [value] happens to be '?', it appears as '.'

# b. They appear as V1, V2, V3, ... etc.
# c. They appear as '?'. Or when colClasses is specified, they cause an error.
# d. Because the default is to store non-numeric values as factors.
# Factors only store the unique values as characters (and the specific value 
# taken at an entry is stored as an integer). This is much cheaper in memory
# than storing everything as a character, because many values repeat themselves.

# e. It becomes a matrix i.e. a vector of vector Each vector at the first level
# is a column of the dataframe, at the second level contains the entries in
# that column.

# 4)
```{r}
hist(auto$price, col='green')
boxplot(auto$horsepower, horizontal=TRUE)

tmp <- table(auto$body_styles)
prop.table((tmp))
tmp <- tmp[order(tmp, decreasing=TRUE)] 
barplot(tmp)

stars(auto[auto$aspiration=='turbo',
           c('wheel_base', 'length', 'width', 'height', 'price')])
```

# 5)
```{r}
# a)
mean(auto[auto$fuel_type=='gas','price'], na.rm=TRUE)
# b)
auto[auto$num_of_cylinders=='twelve', 'make']
# c)
tmp <- table(auto[auto$fuel_type=='diesel', 'make'])
names(tmp)[which.max(tmp)]
# d)
auto[which.max(auto$horsepower), 'price']
# e)
quantile(auto$city_mpg, 0.1)
# f)
quantile(auto$highway_mpg, 0.9)
# g)
quantile(auto$city_mpg, 0.05)
```

# 6)
# a) Returns NULL
# b) 4
# c) Because mpg is not a variable. "mpg" instead creates a string.
# d) Not in general, unless the list has the same length as the number of rows
# i.e. the following code will run, and each value in the column xyz will be assigned
# the corresponding entry in the list
# tmp <- as.list(1:204)
# auto$xyz <- tmp
# e) It returns a list of vectors, each vector contains the values in that column.
# f) data.frame()

# 7)
```{r}
qdat <- na.omit(auto[, c(-1,-2)])
qdat <- qdat[, sapply(qdat, is.numeric)]
cor(qdat)
```
# 8.1)
```{r}
pca_result <- prcomp(qdat, scale.=TRUE)
var = pca_result[1]$sdev ^ 2
sum(var[1:3]) / sum(var)
```

# 8.2)
```{r}
Pframe <- pca_result[2]$rotation[,1:2]
P <- data.matrix(Pframe)
plot(data.matrix(qdat) %*% P)
plot(Pframe)
biplot(pca_result)
```

# Based on the first plot, most variations are captured by the first two principle
# components (because scatters mostly lie on a line). That is, roughly speaking the 
# data is linearly explainable and effective compression can be achieved.
# Most vehicles tend to have low values on the first two principle components.
